---
title: "Object Initialization and preprocessing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Object Initialization and preprocessing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}


options(crayon.enabled = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, warning=FALSE, message=FALSE}
library(Rcongas)
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyverse)
```

Setup the directory where results are stored.
```{r}

sample = 'lymphoma'

prefix = paste0(sample, "/tutorial/")

out.dir = paste0(prefix, "/congas_data/")
fig.dir = paste0(prefix, "/congas_figures/")

if (!dir.exists(out.dir)) {dir.create(out.dir, recursive = T)}
if (!dir.exists(fig.dir)) {dir.create(fig.dir, recursive = T)}
```

Load the ATAC counts to create the tibble needed as CONGAS+ input
```{r}

data(atac_counts) 
data(rna_counts)

```

With the multiome assay, the feature files provided by CellRanger contains already
a mapping between gene names and coordinates. Let's read this file
As an alternative, when such file not not readily available, users can exploit
BiomaRt Please see the correspodning vignette \code{Gene mapping}

```{r}
data(features)
```


Create the RNA and ATAC tibble required by CONGAS+
```{r}
atac_featureDF = data.frame(id = rownames(atac_counts)) %>%
  tidyr::separate(id, c('chr', 'from', 'to'), '-')

atac = Rcongas::create_congas_tibble(counts = atac_counts, 
                                     modality = 'ATAC', 
                                     save_dir = NULL,
                                     features = atac_featureDF)
# duplicate call names across ATAC and RNA cells is not allowed by CONGAS+
# add suffix to atac cell barcodes prior to create the CONGAS+ object
atac = atac %>% mutate(cell = paste0(cell, '-ATAC'))

# Compute normalization factors
norm_atac = Rcongas:::auto_normalisation_factor(atac) %>%
  mutate(modality = 'ATAC')

rna = Rcongas::create_congas_tibble(counts = rna_counts, 
                                    modality = 'RNA', 
                                    save_dir=NULL, 
                                    features = features)

rna = rna %>% mutate(cell = paste0(cell, '-RNA'))

# Compute normalization factors
norm_rna = Rcongas:::auto_normalisation_factor(rna) %>%
  mutate(modality = 'RNA')

atac = atac %>% mutate(value = as.integer(value))
rna = rna %>% mutate(value = as.integer(value))
```

Apply some basic filters on known genes
```{R}
rna = rna %>% filter_known_genes(what='r')

all_genes = rna$gene %>% unique
mito = all_genes %>% str_starts(pattern = 'MT-')
all_genes = setdiff(all_genes, all_genes[mito])
rna = rna %>% dplyr::filter(gene %in% all_genes)
```

Read chromosome arm coordinated that will be used as segment breakpoints
```{R}
data(example_segments)

# Remove these chromosomes
segments = example_segments %>% dplyr::filter(chr != 'chrX', chr != 'chrY')
```

Now init the CONGAS+ object

```{R}
x = init(
  rna = rna,
  atac = atac,
  segmentation = segments, #%>% mutate(copies = as.integer(round(copies))),
  rna_normalisation_factors = norm_rna,
  atac_normalisation_factors = norm_atac,
  rna_likelihood = "NB", 
  atac_likelihood = "NB",
  description = paste0('Bimodal ', sample))
```

Now plot the number of nonzero cells vs the number of peaks per segment, to help 
identify a threshold for filtering segments. 

```{R}
ggplot(x$input$segmentation, aes(x=ATAC_nonzerocells, y=ATAC_peaks)) + geom_point() + theme_bw()
ggplot(x$input$segmentation, aes(x=RNA_nonzerocells, y=RNA_genes)) + geom_point() + theme_bw()
```

```{R}
x = Rcongas::filter_segments(x, RNA_nonzerocells = 300, ATAC_nonzerocells = 300)
```

Generate plots to check out the data distribution
```{r, fig.width=8, fig.height=8}
plot_data(
  x,
  what = 'histogram',
  segments = get_input(x, what = 'segmentation') %>%
    ungroup() %>%
    mutate(L = to - from) %>%
    dplyr::arrange(dplyr::desc(L)) %>%
    dplyr::top_n(9) %>%
    pull(segment_id)
)
```


Add metadata to annotate the counts distributions
```{r, fig.width=8, fig.height=8}
data(metadata)

x$input$metadata = metadata
# Modificato plot_data in mdodo da plottare gli istogrammi colorati per celltype
plot_data(x, 
  to_plot = 'type', 
  position = 'stack',
  segments = get_input(x, what = 'segmentation') %>%
    ungroup() %>%
    mutate(L = to - from) %>%
    dplyr::arrange(dplyr::desc(L)) %>%
    dplyr::top_n(9) %>%
    pull(segment_id))
```

Please see the next vignette \code{Fitting} that shows how to perform CONGAS+ inference.