---
title: "Fitting a model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting a model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}


options(crayon.enabled = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, warning=FALSE, message=FALSE}
library(Rcongas)
library(dplyr)
library(ggplot2)
library(stringr)
```

Setup the directory where results are stored.
```{r}

sample = 'lymphoma'

prefix = paste0(sample, "/tutorial/")

out.dir = paste0(prefix, "/congas_data/")
fig.dir = paste0(prefix, "/congas_figures/")

if (!dir.exists(out.dir)) {dir.create(out.dir, recursive = T)}
if (!dir.exists(fig.dir)) {dir.create(fig.dir, recursive = T)}
```

We load the object created in the previous Vignette \code{create_congas_object}.
```{r}

data(multiome_congas_object) 

```

To reduce the paramter search space dimensions and subsequently decreasing the computational complecity of CONGAS+ inference, users can optionally decide to perform 
a preliminary step of segment filtering, which consists in running CONGAS+ inference independently on each segment, varying the number of clusters from 1 to 3 and
finally keeping only those segments in which the otpimal number of clusters selected by BIC is higher than 1. 

This step is implemented in the fuction `segments_selector_congas`, which return the filtered CONGAS+ object.

Fitting uses `reticulate` to interface with the Python CONGAS package, which implements the models in Pyro. In case R does not find a anaconda environment with CONGAS+
python version installed, it will automatically create a r-reticulate environment and install CONGAS+ within that environment.
```{r, results = 'hide', message=FALSE}

filt = segments_selector_congas(multiome_congas_object)
# You can save the filtering result to avoid re-running the whole pipeline in future steps
# saveRDS(filt, paste0(out.dir, "rcongas_obj_filtered.rds"))
```

Now we can run CONGAS+ parameters inference on the final filtered object. We first  using \code{auto_config_run}, and then we fit the model.
The function \code{auto_config_run} computes the hyperparameters values using the current data. 
Its parameter \code{normal_cells} is used to specify if the data contains a subset of normal cells. 
In case this is set to \code{TRUE}, the copy number distribution for one of the clusters will be initialized with values skewed towards the diploid state in every segment.
```{r}
# Set values fro the model hyperparameters
k = c(1:4)

binom_limits = c(40,1000)
model = "BIC"

lr = 0.01
temperature = 20

steps = 5000
lambda = 0.5

# Estimate hyperparameters
hyperparams_filt <- auto_config_run(filt, k, 
                                    prior_cn=c(0.2, 0.6, 0.1, 0.05, 0.05),
                                    init_importance = 0.6, 
                                    CUDA = FALSE, 
                                    normal_cells=TRUE)

# Run
hyperparams_filt$binom_prior_limits = binom_limits

fit_filt <- Rcongas:::fit_congas(filt,
                                 K = k, 
                                 lambdas = lambda, 
                                 learning_rate = lr, 
                                 steps = steps,
                                 model_parameters = hyperparams_filt, 
                                 model_selection = model,
                                 latent_variables = "G",
                                 CUDA = FALSE,
                                 temperature = temperature, 
                                 same_mixing = TRUE, 
                                 threshold = 0.001)
```

The new object has more information.
```{r}
fit_filt
```
You can get the information regarding model selection metrics
```{r}
fit_filt$model_selection
```

```{r, fig.width=5, fig.height=4}
Rcongas::plot_fit(fit_filt, what='scores')
```

```{r, fig.width=5, fig.height=8}
plot_fit(fit_filt, 'posterior_CNA')
```

```{r, fig.width=20, fig.height=10}
p = cowplot::plot_grid( 
  plotlist = plot_fit(fit_filt, what = 'density', highlights = TRUE),
  ncol = 4)
p
ggsave(filename= paste0(fig.dir, 'clusters_distribution.png'), plot= p, height = 15, width = 15)
```

