---
title: "Filtering functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Filtering functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(crayon.enabled = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# old_hooks <- fansi::set_knit_hooks(knitr::knit_hooks, 
#                                    which = c("output", "message", "error"))

```

```{r, warning=FALSE, message=FALSE}
library(Rcongas)
library(dplyr)
```


```{r}
data('example_object')

print(example_object)
```

# Filtering functions 

Filtering functions are available for data _before_ it is loaded into the package, or _after_.

## Before loading

Functions that pre-process data before being inputed to the tool work requiring data in a format ready for usage in the tool; the output is the very same tibble with filtered entries (rows).

## After loading

Functions that process data that has already being inputed to the tool work and return an \code{rcongasplus} object.

### Removing outliers from segments

After mapping counts data to segments, a function
can be used to determine quantiles of mapped data, and identify
outliers in each segment and modality. 

An outlier can then be removed or capped to the median cell value.
The former option introduced 0-counts in the data, which we suggest
to check with the `stat` function, and possibly remove by using
the `filter_missing_data` function. Removal can be important as
an excess of 0-counts cells (_missing data_) will drive the fit
to use 0-mean components.

Outliers are identified by using a quantile range $q = (q_l, q_r)$ with 
$q_r,q_l\in [0,1]$ reporting lower and upper quantiles. The quantiles 
are computed and applied per segment. 

```{r fig.width=8, fig.height=7}
# 1 segment from the object
segment_id = example_object %>% 
  get_input('segmentation') %>% 
  filter(row_number() == 1) %>% 
  pull(segment_id)
  
# Original segment plot
segment_plot = plot_data(example_object, what = 'histogram', segments = segment_id)

# Filter with default values, capping
f_example_object = example_object %>% filter_outliers(action = 'cap')

# Capped segment plot
segment_plot_2 = plot_data(f_example_object, what = 'histogram', segments = segment_id)

# Filter with default values, remove entries
f2_example_object = example_object %>% filter_outliers(action = 'remove')

# Removed segment plot
segment_plot_3 = plot_data(f2_example_object, what = 'histogram', segments = segment_id)

# Show the difference
cowplot::plot_grid(
  segment_plot,
  segment_plot_2, # capped
  segment_plot_3, # removed
  nrow = 3,
  ncol = 1
)
```


> If you used \code{action = 'remove'} The returned object will contain fewer entries; it should be checked for an excess of zeros after filtering. We show an example below

Compare 0-count cells before and after the filter

```{r}
# RNA 0-counts difference
stat(example_object)$zero_counts_cells_RNA %>% print
stat(f_example_object)$zero_counts_cells_RNA %>% print  # capped
stat(f2_example_object)$zero_counts_cells_RNA %>% print # removed
```

```{r}
# ATAC 0-counts difference
stat(example_object)$zero_counts_cells_ATAC %>% print
stat(f_example_object)$zero_counts_cells_ATAC %>% print  # capped
stat(f2_example_object)$zero_counts_cells_ATAC %>% print # removed
```

The extent of the new 0-counts cells can be summarised printing the filtered object to console.

```{r}
f_example_object %>% print
f2_example_object %>% print
```

### Removing 0-count cells

Some cells can have have 0-counts because they have no events mapping to a certain segment, or because other filtering functions have removed entries, an inflated the number of 0-count cells.

A filter is available to remove cells that have 0-counts in a proportion $p\in[0,1]$ of 

> 0-count entries will be imputed to have value 0 by the fit algorithm. In case of an excess of 0-counts the model will try to explain 0-values with mixture components peaked at 0, which can make the inference more instable. We suggest therefore to remove such entries with this filtering function.

We make a concrete example working with the `f2_example_object` object creted by the previous filter. As shown above, after filtering segment's outliers 

```{r}
# Observe the number of cells
print(f2_example_object)

# Apply the filter
f2_example_object = f2_example_object %>% filter_missing_data()

# Reduced number of cells
print(f2_example_object)
```

